---
title: "Data Import with Readr"
author: "Fernando Figueres"
subtitle: "DATA-607, Spring 2019"
output:
  html_document:
    df_print: paged
---

# Introduction

Readr's primary function is to load plain-text rectangular files into R data frames. It's also able to write data frames to different flat files and it includes additional functions to address different data types, unconventional delimiters, encoding, etc.

In this vignette, we will explain how to install and load the package and then explore its basic functionality by loading a CSV file into R. Finally, we will save our data frame as a TSV file.

By default, Readr will try to assign the appropriate data type to each column. In our example with will import a subset of the columns using `cols_only` and we'll use the `col_*` functions to specify each data type.

# Installation

Readr can be installed along with the complete set of Tidyverse packages by using the command:

```{r eval=FALSE}
install.packages("tidyverse")
```

Alternativley, it can be installed by itself by running:

```{r eval=FALSE}
install.packages("readr")
```

# Loading and Basic Use

Before using the package it should be loaded. As before, it can be loaded along with the rest of the Tydiverse packages using the command below:

```{r}
library(tidyverse)
```

Or it can be loaded independently using:

```{r}
library(readr)
```

Readr is designed to import 7 types of flat files and has a specific function for each:

read_csv(): comma separated (CSV) files  
read_tsv(): tab separated files  
read_delim(): general delimited files  
read_fwf(): fixed width files  
read_table(): tabular files where columns are separated by white-space.  
read_log(): web log files  

By default, Readr will try to assign the appropiate data type to each column. 

col_logical(): containing only T, F, TRUE or FALSE.  
col_integer(): integers.  
col_double(): doubles.  
col_character(): strings, and everything else.  
col_factor(levels, ordered):a fixed set of values.  
col_date(format = ""): with the locale's date_format.  
col_time(format = ""): with the locale's time_format.  
col_datetime(format = ""): ISO8601 date times.  
col_number() numbers containing the grouping_mark  .
col_skip(): don't import this column.  
col_guess(): parse using the "best" type based on the input.  

## Basic import with default options

Lets import our CSV file using Readr's default options.

```{r}
fifa <- read_csv(
  file = "fifa.csv")
```
As shown above, Readr will output the result of the operation along with the data type for each variable. In some cases, the output is truncated so we'll use the `spec()` function to get the full list.

```{r}
spec(fifa)
```

There are more variables than we need and many of them have a data type which isnt' optimal. Next we'll import a subset of variables and manually assign data types.

## Filtered import

```{r}
fifa_subset <- read_csv(
  file = "fifa.csv",
  cols_only(
    Name = col_character(),
    Age = col_integer(),
    Nationality = col_factor(),
    Club = col_factor(),
    Joined = col_date(format = "%b %d, %Y")),
  col_names = TRUE
  )
```

We now have a data frame with our 5 variables, each with an appropiate data type.

```{r}
head(fifa_subset)
```

# Saving data frames to file

Perhaps we have finished with out analysis and we'd like to share our data with a colleague. We can save our data to a file using Readr's `write_*` functions.

In this case, we can save it as a tab delimted file by using the following code.

```{r}
write_tsv(
  fifa_subset,
  path = "fifa_subset.tsv"
)
```



**References**

https://readr.tidyverse.org/

https://readr.tidyverse.org/reference/cols.html

https://www.kaggle.com/karangadiya/fifa19

https://www.ibm.com/support/knowledgecenter/en/SSQHWE_10.1.0/com.ibm.ondemand.mp.doc/arsa0257.htm

